---
title: "Lab 9: Clustering"
subtitle: "Lab 9"
editor: visual
execute: 
  message: false
  warning: false
---

## Lab 9 - Clustering

## Instructions:

-   If you are fitting a model, display the model output in a neatly formatted table. (The \`tidy\` and \`kable\` functions can help!).

-   If you are creating a plot, use clear labels for all axes, titles, etc.

-   Commit and push your work to GitHub regularly, at least after each exercise. Write short and informative commit messages.

When you're done, we should be able to knit the final version of the QMD in your GitHub as a HTML.

## Data

-   For this lab, we will reproduce the results of Reilly et al. (2023; Experiment 1). Do not worry about any of the statistical tests performed.

Reilly, J., Finley, A. M., Kelly, A., Zuckerman, B., & Flurie, M. (2021). Olfactory language and semantic processing in anosmia: A neuropsychological case control study. *Neurocase*, *27*(1), 86--96. <https://doi.org/10.1080/13554794.2020.1871491>

-   The data and analysis scripts can be found on OSF.

## Hierarchical clustering

-   Run a hierarchical cluster analysis as done in Reilly et al. (2021). Please include the both dendrograms. Do your dendrograms have similar pattern?

    ```{r}

    ```

-   Run the same analysis, but choose a different link method. Does this change the interpretation at all? Please provide a dendrogram with the link method you selected.

    ```{r}

    ```

-   Determine the optimal number of clusters for each group using the elbow, silhouette, and gap methods. Do they all agree?

    ```{r}

    ```

-   Perform a bootstrapped hierarchical clustering analysis on both P01 and MTurk samples. Does the number of clusters converge with the \# clusters above?

    ```{r}


    ```

-   Include a color-coded dendrogram for your final selection (use `dendextend`)

## K-means

-   Now perform a k-means cluster analysis. As a starting point, k = \# of clusters decided on above. Plot a figure of the cluster solution.

    ```{r}

    ```

-   Use the consensus method `n_clusters` to determine the number of clusters. Include a figure of the new cluster solution.

    ```{r}

    ```

    ## Write-up

-   Write-up a report and make sure to interpret/name your clusters. If there are any differences between your analysis and theirs, describe them. Additionally, share your thoughts on how the authors shared their data.
